### 2017-04-28 paper.tex
start writing paper

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\graphicspath{{images/}}

\title{The User Behavior On Online Judge Website}
\author{yueyi zhuo}
%\institute{School of Mathematical,Sichuan Normal University}
\date{2017}

\begin{document}

\maketitle

\begin{abstract}

Abstract content.

\end{abstract}

\section{Introduction}

In 2016 Spring Festival, I solve some problems in Project Euler 
\footnote{In fact, I solve 70+ problems before my interest fade.}, 
which may be most famous online judge 
in Internet except Leetcode. I have noticed the interesting data structure related with rank and difficulty.
But I lacked enough powerful statistic tool to process and model it. 
In fact, I model it as a ordinary least squard which is not understood by me in the time. In the naive period,
linear regression model seems like a poor black box model that often be meeted in Machine Learning algorithm.

\section{Motivation}

Motivation content.

\end{document}